%---------change this every homework
% place your email id between the braces so that your homework has a name
\def\yourname{Hamlet}
% -----------------------------------------------------
\def\homework{2} % 0 for solution, 1 for problem-set only
\def\duedate{Tuesday, January 26 at 12 noon}
\def\duelocation{via \href{https://www.gradescope.com/courses/80616/}{gradescope}}
\def\hnumber{2}
\def\prof{Andrew Drucker, Lorenzo Orecchia}
\def\course{\href{https://canvas.uchicago.edu/courses/25977}{CMSC 27200 - Winter 2021}}
%-------------------------------------

\documentclass[10pt]{article}
\usepackage[colorlinks,urlcolor=blue]{hyperref}
\usepackage[osf]{mathpazo}
\usepackage{amsmath,amsfonts,graphicx}
\usepackage{latexsym}
\usepackage{subfig}
\usepackage{algpseudocode}
%\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{listings}
%\usepackage[top=1in,bottom=1.4in,left=1.5in,right=1.5in,centering]{geometry}
\usepackage{fullpage}
\usepackage{color}
\definecolor{mdb}{rgb}{0.3,0.02,0.02} 
\definecolor{cit}{rgb}{0.05,0.2,0.45} 
%\pagestyle{myheadings}
\markboth{\yourname}{\yourname}

\thispagestyle{empty}

\newenvironment{proof}{\par\noindent{\it Proof.}\hspace*{1em}}{$\Box$\bigskip}
\newcommand{\qed}{$\Box$}
\newcommand{\alg}[1]{\mathsf{#1}}
\newcommand{\handout}{
   \renewcommand{\thepage}{H\hnumber-\arabic{page}}
   \noindent
   \begin{center}
      \vbox{
    \hbox to \columnwidth {\sc{\course} --- \prof \hfill}
    \vspace{-2mm}
    \hbox to \columnwidth {\sc due \MakeLowercase{\duedate} \duelocation\hfill {\Huge\color{mdb}H\hnumber.\yourname}}
      }
   \end{center}
   \vspace*{2mm}
}
\newcommand{\solution}[1]{\medskip\noindent{\color{cit}\textbf{Solution:} #1}}

\newcommand{\bit}[1]{\{0,1\}^{ #1 }}
%\dontprintsemicolon
%\linesnumbered
\newtheorem{problem}{\sc\color{cit}problem}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{claim}{Claim}


\begin{document}
\handout
\begin{itemize}
\item The assignment is due at Gradescope on \duedate. 

\item You can either type your homework using LaTex or scan your handwritten work. We will provide a LaTex template for each homework. If you writing by hand, please fill in the solutions in this template, inserting additional sheets as necessary. This will facilitate the grading.

\item You are permitted to study and discuss the problems with 2 other students (per problem; any section).  However, {\em you must write up your own solutions, in your own words}. Do not submit anything you cannot explain. If you do collaborate with any of the other students on any problem, please do list all your collaborators in your submission for each problem. 

\item Similarly, please list any other source you have used for each problem, including other textbooks or websites.


\item {\em Show your work.} Answers without justification will be given little credit.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{problem}[25 points]
Solve exercise 3 in Chapter 4 in the Kleinberg-Tardos textbook.  (trucking)
\end{problem}

\solution{
  Collaborated with: Yael Sulkin, Bryan Lee \\

\paragraph{Algorithm} The greedy algorithm this proof will be proving is the following:
      
\begin{algorithmic}
      \State Let all trucks have maximum weight, $W$ 
      \State Let the number of trucks sent to Boston, $numTrucks$. 
      \State Let the truckweight at $k$th iteration be $truckWeight$. 
      \State Let $i$ be the package that most recently arrived, woth weight of i be $w_i$ \\

      \While {packages arrive}  
        \If {$truckWeight + w_i < W$} 
          \State update $truckweight = truckweight + w_i$
        \Else 
          \State increase $numTrucks$ by 1
          \State reset $truckWeight$ to $0$
        \EndIf
      \EndWhile
\end{algorithmic}

\paragraph{Proof} We will prove this algorithm inductively. For our greedy algorithm we want to show that it "stays ahead" or any other solution
such that at the $k$th iteration our algorithm is at most the same as the optimal solution. Let $A$ represent the number of trucks used by our greedy algorithm, 
$O$ represent the number of trucks used by the optimal solution, and let $A_k$, $O_k$ represent the respective quanity of trucks used at any given iteration $k$. 
Four our base case, we set $k = 1$. When our first package arrives to the station it is loaded on to the first truck in both algorithms. We know that for packagae
$i$, $w_i \leq W$ which means the number of trucks used is at most $1$. This solved our base case. Now, let $k > 1$ and assume our greedy algorithm stays ahead for 
$j < k$ iterations. We want to show that when $A_k \leq O_k$ at any given iteration then $A_{k+1} \leq O_{k+1}$. That is our inductive hypothesis and we can prove it 
using case analysis. We can break the cases down to $A_{k} = A_{k+1}$ and $A_{k} < A_{k+1}$. First if we know that $A_k \leq O_k \leq O_{k+1}$ and $A_{k} = A_{k+1}$
then we know $A_{k+1} \leq O_{k+1}$ which proves our inductive hypothesis for this case. Secondly, if $A_k < A_{k+1}$ ... not finished
}


\newpage
\medskip\noindent{\color{cit} Extra Space for your solution}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{problem}[25 points]
Solve exercise 5 in Chapter 4 in the Kleinberg-Tardos textbook.  (cell phone towers)

Please note!  Here and elsewhere, when the authors (or your instructors) say ``give an algorithm'' without further instructions, they mean 
``give an algorithm AND prove that it is correct and runs in polynomial time''.  This should be assumed henceforth.
\end{problem}
\solution{
  Collaborated with: Yael Sulkin, Bryan Lee\\

  \begin{algorithmic}
    \State Let $h$ be in set $H$ of all houses along the road.
    \State Let $x_h$ be the locatin of house $h$ where $x_h$ = 0 if at westernmost point and the easternmostpoint is at $x_h$ max. 
    \State Let $B$ represent the set of base towers placed

    \While {$h \in H$}
      \If {$x_h + 4 \geq$ easternmost point}
        \State place base on easternmost point
        \State add base to set B
        \State delete all houses in $H$ s.t. $x_4 + 4 \geq$ easternmost point
      \Else 
        \State place a base in $x_h + 4$
        \State add base to set B
        \State delete all houses in H s.t. $x_h' \leq x_h + 8 miles$
      \EndIf
    \EndWhile

  \end{algorithmic} 
 
  \paragraph{}We know our algorithm runs on polynomial time because our loop iterates over the set of houses once, returning the set of bases after reaching the last house. 
  Our algorithm runs on $O(n)$ time by definition. 

  \paragraph{Proof} Let set $A$ be the set of bases placed by our algorithm and set $O$ be the set of bases placed 
  by the optimal algorithm. If our set $A$ is not optimal and we are looking for the minimal amount of bases then 
  we know that if $|A| = n$ and $|O| = m$ then $m < n$. If $m < n$ then we know there exists some base, $b$ in $A$ such that
  it either covers $0$ houses or there also exists some $b'$ such that $b'$ can be moved to cover the houses within $b$. We will continue this 
  proof using case analysis and proof by contradiction. 

  \paragraph{} In the case where $b$ in $A$ covers $0$ houses, this, in fact, contradicts the algorithm for $A$ which states 
  "while there exists house $h \in H$". This case is impossible for our proposed greedy algorithm. 

  \paragraph{} In the case where $b'$ exists in $A$ such that $b'$ can be moved, let $b$ be located at distance $x$ and $b'$ be located at distance $x'$ 
  such that $x \neq x'$ and $x, x' \geq 0$. According to our case statement, there exists $h$ in ${x - 4, x + 4}$ such that $b'$ can be moved to cover 
  ${x - 4, x + 4} \cup {x' - 4, x' + 4}$ or ${x - 4, x + 4, x' - 4, x' + 4}$. $b'$ will have to at most cover 
  $|(x - 4) - (x' + 4)| = |x - x' - 8|$.
  If the distance $b'$ has to cover is greater than 8 this contradicts our givens. If $|x - x' - 8| = 0$ this means base $b$ and base $b'$ are exactly 8 meters apart. Let $b$ be west of $b'$. 
  This implies there exists some $h$ such that $x_h$ is exactly halfway between $b$ and $b'$. However, this is impossible because according to our algorithm $b'$ was placed 4 miles east of house $h \in H$ yet
  $h$ was not in $H$ at that iteration because $x_h$ was covered by base $b$ and therefore deleted from available input. 
}


\newpage
\medskip\noindent{\color{cit} Extra Space for your solution}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{problem}[25 points]
The Running Sums problem is defined as follows:

\vspace{1 em}


\textbf{Input:} a sequence $(a_1, \ldots, a_n)$, where each $a_i$ is either 1 or -1.  

\vspace{1 em}

\textbf{Desired output:} a sequence $(b_1, \ldots, b_n)$, where each $b_i$ is either 0 or 1.  

Your goal is to \textbf{minimize} $\sum_{1 \leq i \leq n} b_i$, subject to the \textbf{constraint} that

\[    \sum_{1 \leq i \leq j} (a_i + b_i) \ \geq \ 0 \ ,   \ \ \ \ \ \ \ \ \   \mathrm{for \ all \ \ \ }j \in \{1, 2, \ldots, n\} \ .  \]

Give an algorithm for this problem that, for full credit, should run in $O(n)$ steps.



\end{problem}
\solution{
Collaborated with: Yael Sulkin and Bryan Lee

\begin{algorithmic}
  \State Let $b$ start off as $0$
  \State Let $J$ represent the running sum of $a_i$ and $b_i$
  \State Let $B$ represent the running sum of $b_i$ 

  \While {$a_i$ in $a_1 \ldots a_n$}
    \If {$J > 0$}
      \State $b_i$ remains $0$
      \State Update $J$ and add $a_i$ and $b_i$ 
      \State Keep $B$ the same
    \Else 
      \State $b_i$ can be derived by the equation $b_i = \frac{a_i}{2} + \frac{1}{2}$
      \State Update $J$ by add $a_i$ and $b_i$
      \State Update $B$ by adding $b_i$
    \EndIf
  \EndWhile

\end{algorithmic}

\paragraph{} This algorithm runs on O(n) time as it loops through sequence a once while building towards b at every step. 

\paragraph{Proof} An optimal algorithm increases $b$ minimally at each step and we know $b$ can either be 0 or 1 so our greedy algorithm's goal is to maximize the number 
of $0$'s in b and limit the number of 1's in b. As such, we want more cases where b should be $0$ and less otherwise. Thus, we will prove our greedy algorithm with contradiction. Let 
$A$ represent the sum of $b$'s of my algorithm and $O$ represent the sum of $b'$ of the optimal solution. We want to show $A \leq O$ at any given iteration k. Let $A_k$, $O_k$ 
represent the respective sum of b at kth iteration. We will treat each k iteration using case analysis, particularly looking at why b would ever want to be 1. The only scenario would be if 
$a_i = -1$ and the current running sum at kth iteration of $a_i + b_i = 0$. In this scenario, $b$ can not be 0 or our running sums constraint would be broken. However, let $O \leq A$ then we know 
there exists at least one b in which at some kth iteration, given some $a_k$, $b$ should be 0 and not 1, such that $O_{k+1} increases by 0$. Well if $a_i + b_i = 0$ and $O_{k+1} = 0$ then we know 
$a_i was 1$ which for our algorithm always produces a value of $0$. -not finished
}

\newpage
\medskip\noindent{\color{cit} Extra Space for your solution}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{problem}[25 points]
In the Hopping Game, there is a sequence of $n$ spaces.  You begin at space 0 and at each step, you can hop 1,2,3, or 4 spaces forward. However, some of the spaces have obstacles and if you land on an obstacle, you lose. 

Give a greedy algorithm which, given an array $A[1, \ldots, n-1]$ of Boolean values with $A[i]$ indicating the presence/absence of obstacle at position $i \in [1, n-1]$, find the minimum number of hops needed to reach space $n$ without losing, if it is possible to do so.  (We assume spaces 0 and $n$ are obstacle-free, and are not part of the input.)  Prove that your algorithm is correct. For full credit, your algorithm should run in time $O(n)$.
\end{problem}

\solution{ 
  Collaborated with: Yael Sulkin

\begin{algorithmic}
  \While {for every input $a_i$ in $A$}
    \If {$a_{i + 4}$ is $0$}
      \State add to hops counter
      \State change location to $a_{i + 4}$
    \Else 
      \If {$a_{i + 3}$ is $0$}
        \State add to hops counter
        \State change location to $a_{i + 3}$
      \Else 
        \If {$a_{i + 2}$ is $0$}
          \State add to hops counter
          \State change location to $a_{i + 2}$
        \Else 
          \If {$a_{i + 1}$ is $0$}
            \State add to hops counter
            \State change location to $a_{i + 1}$
          \Else 
            \State Return hops counter, Terminate loop. Reaching N is not possible
          \EndIf
        \EndIf
      \EndIf
    \EndIf
  \EndWhile 


\end{algorithmic}

\paragraph{} This algorithm has runtime O(n) since we work our way through array $A$ from left to right once
and never loop over any square multiple times. Our algorithm's runtime is proportional to the size of $A$ which by definition is 
O(n). 

\paragraph{Proof} Let the number of hops made by our algorithm be denoted by $H$ and by the most optimal algorithm be denoted by $O$. 
We will prove our algorithm provides at most the same amount of hops as the optimal solution using proof by induction. We will denote 
the number of hops at iteration, $k$, made by our greedy algorithm as $H_k$ and those by the optimal algorithm as $O_k$. Our base case
is when $k = 1$. When $k = 1$, both algorithms are starting from space $0$. The optimal algorithm will jump once if there is indeed a 
path to space $n$, thus, $O_1 = 1$. Similarly, our greedy algorithm will jump at least some distance if there is an avaialable square 
within the next 4 tiles thus $H_1 = 1$. Our base case is proved. We must show that for the $k+1$st case our greedy algorithm stays ahead
and at each iteration $H \leq O$. Well, for each hop in the optimal algorithm, the player gets closest to $n$ such that the distance between
the player and space $n$ is minimal after each hop. We want to show our greedy algorithm minimizes the distance to space $n$ with each hop. 
For the optimal algorithm we know 1 hop = n - (hop distance + $a_i$) where $n - (hop distance + a_i)$ is minimized. Since, $n$ is a constant and 
$a_i$ is also a constant representing current location, we know that our greedy algorithm must maximize hop distance at each iteration. Our greedy 
algorithm does, in fact, by assuming maximum hop distance and working our way to smaller distances only if greater ones aren't available.

}


\newpage
\medskip\noindent{\color{cit} Extra Space for your solution}



\end{document}
